---
title: "Introduction to the proteintools package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{proteintools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(proteintools)
```

The proteintools package provides a set of tools for basic analysis of protein and DNA sequences. It is intended for educational and exploratory use in bioinformatics. This package was developed as part of a data science for biology 2 portfolio project at Hogeschool Utrecht.

The package currently includes the following functions:

`protein_mass()` – Calculate the molecular mass of a protein sequence.     
`is_valid_protein()` –  Check if a sequence contains only valid amino acids.     
`analyze_amino_acid()` – Analyze amino acid in protein sequence.     
`translate_dna()` – Translate a DNA sequence to a protein.      
      
## Protein mass

You can calculate the mass of a protein sequence using the protein_mass() function.

```{r}

protein_mass("MKT")

```

## Valid protein

Check whether a sequence contains only valid amino acids.

```{r}

is_valid_protein("MKTWSSYF")
is_valid_protein("X")

```


## Analyse amino acids

Count the number of a provided amino acid and returns the location(s).

```{r}

analyze_amino_acid("MKTWSSYF", "S")

```


## DNA Translation

Translate a DNA sequence to a protein.

```{r}

translate_dna("ATGGCCATTGTAATGGGCCGCTGAAAGGGTGCCCGATAG")

```
       
       
       
# Conclusion

This vignette introduced the basics of the proteintools package. The package is designed to make protein related tasks easier. For more details, see the GitHub repository.
